Towards a Mechanized Proof of Noninterference

### 2013-12-18 paper/paper.tex
Added supporting text to both the type system and the proofs.  Still need to give a basic outline of types in the type system section.

At a high level, these bracketed expressions are evaluated by evaluating one of
the branches, as shown in rules \ref{eq:bracket-left} and
\ref{eq:bracket-right}, or by expanding the brackets around the parent
expression, passing the appropriate branch of the bracketed subvalue to each
branch of the resulting expression, as shown in rules \ref{eq:lift-app} and
\ref{eq:lift-if}.  In these rules, we use the notation $\lift{x}_i$ to denote
the left ($i = 1$) or right ($i = 2$) branch of a bracketed expression or value.

@@

The language's type system includes subtyping, which allows data that have a
lower label to flow into higher labelled data.  The rules for subtyping are
given in Figure~\ref{fig:subtype_rules}.  Rule \textsc{S-Int} states that
integers are subtypes of integers with equivalent or higher labels in the label
lattice.  Similarly, rule \textsc{S-Abs} gives the standard subtyping
relationship for functions along with the same label relation.

@@

For the sake of being able to discuss how strong labels are, we follow the
example of Simonet and Pottier~\cite{InfoFlowML} and define a ``guards''
relation between labels and types, denoted $l \guards t$, which states that the
label $l$ is at least as restrictive as the label on the type $t$.  In other
words, this requires that the type has at least $l$ for it's security label.
The two rules regarding this relationships for both types in our language are
given in Figure~\ref{fig:guard_rules}.

@@

Figure~\ref{fig:value_types} includes typing judgements for values and
Figure~\ref{fig:expr_types} includes typing judgements for expressions.

@@

@@

\textbf{TODO: Talk more about the rules and what they're doing.  Specifically,
discuss what the heck H and pc are doing.}

@@

\textbf{TODO: Add in text between lemmas and theorems discussing what each one
is at a high level.}

@@

@@

@@

@@

@@

@@

>>>>

At a high level, these bracketed expressions are evaluated by evaluating one of
the branches, as shown in rules \ref{eq:bracket-left} and
\ref{eq:bracket-right}, or by expanding the brackets around the parent
expression, passing the appropriate branch of the bracketed subvalue to each
branch of the resulting expression, as shown in rules \ref{eq:lift-app} and
\ref{eq:lift-if}.  In these rules, we use the notation $\lift{x}_i$ to denote
the left ($i = 1$) or right ($i = 2$) branch of a bracketed expression or value
$x$.

@@

\textbf{TODO: Give definition of types in the language and discuss stuff that
was dropped from \origLang}

The language's type system includes subtyping, which allows data that have a
lower label to flow into higher labelled data.  We use the notatoin $x \leq y$
to denote that $x$ is a subtype of $y$, whether both $x$ and $y$ are labels or
types (as long as both are the same kind of value).  The rules for subtyping are
given in Figure~\ref{fig:subtype_rules}.  Rule \textsc{S-Int} states that
integers are subtypes of integers with equivalent or higher labels in the label
lattice.  Similarly, rule \textsc{S-Abs} gives the standard subtyping
relationship for functions along with the same label relation.

@@

For the sake of being able to discuss how strong labels are in relation to a
type, we follow the example of Simonet and Pottier~\cite{InfoFlowML} and define
a ``guards'' relation between labels and types, denoted $l \guards t$, which
states that the label $l$ is at least as restrictive as the label on the type
$t$.  In other words, this requires that the type has at least $l$ for it's
security label.  The two rules regarding this relationships for both types in
our language are given in Figure~\ref{fig:guard_rules}.

@@

Figure~\ref{fig:value_types} includes typing judgements for values in the
language, which are mostly straight forward with the exception of
\textsc{V-Bracket}, which we discuss below along with the type rules for
expressions.

@@

The typing rules for expressions in the language are given in
Figure~\ref{fig:expr_types}.  Unlike a typical typing judgment, we include an
additional piece of context information, the $pc$, which is a label that
describes what level of label can influence the control structure decisions to
reach the current expression in the program.  This is a standard extension for
languages with information flow types, like Jif.

The interesting rules in the type system are \textsc{V-Bracket},
\textsc{E-Bracket}, and \textsc{E-If}.  \textsc{E-If} is a standard rule for
information flow type systems, stating that the type determinations for each
branch must have a $pc$ equivalent to the security of the original $pc$ along
with the security of the label on the value in the test, $l$.

\textsc{V-Bracket} and \textsc{E-Bracket} both require that there exists a label
from the set $H$ which representes an upward closed sublattice of the labels in
the system representing what is considered to be ``high'' security for the
purposes of the proof of noninterference, discussed later.  These rules require
that the types of each branch of a bracketed value or expression can be typed
with a high security label.  Furthermore, \textsc{E-Bracket} requires that the
$pc$ of the subexpressions must be at least as high as this same label,
requiring that we treat everything in the subexpression as affected by high
security information.

@@

@@

We start with soundness and completeness lemmas for the operational semantics,
showing that this language can accurately describe the process of evaluating two
different runs of a language where bracketed expressions are high security
values.  Specifically, the image of a valid reduction of an expression is also a
valid reduction and if both runs of a program evaluate to an answer, then so can
the bracketed expression in \langName.

@@

Next, we prove lemmas related to the type system.  We start with a lemma showing
that $\guards$ interacts well with subtyping and a weakening lemma, showing that
anything typed at a given $pc$ can be typed at a lower $pc'$ as well.

@@

We now prove a projection lemma, showing that if a top level expression or value
is well typed, so must either projection of that expression or value to one
program run or another.

@@

Below is a lemma showing that if a bracketed expression is well typed, then
there must exist a high security label to use for the types and $pc$ of both
branches.

@@

Here we prove a standard substitution lemma for the type system.

@@

Finally, we prove one final lemma and the noninterference theorem.  The lemma
shows that if an expression is typed with a low security label, then both
projections (so the value in either run of the program) are equivalent.

@@

Finally, the noninteference theorem shows that if we have an expression that is
typed with a low label, then replacing any bracketed value in the expression
with either projection will result in the same value for the overall expression.
In other words, if we vary any high security item in a low security expression,
the value must not be affected.
